---
title: "p8105_hw2_mw3845"
author: "Minghe Wang"
date: "2024-09-29"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1

read csv into dataframe first

```{r problem1_read_and_clean_data}

subway_df = 
  read_csv('./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv', na = "") |>
  janitor::clean_names() |>
  select(line:entry, vending, ada) |>
  mutate(
    entry = ifelse(entry=='YES', TRUE, FALSE)
  )
tail(subway_df)
```

So far, we read and clean the data of NYC's subway. The data contains ` line`, `station_name`, `station_latitude`, `station longitude`, `route1` to `route11`, `entry`, `vending`, `entrance_type`, and `ada`. 
- Cleaning steps: the variables' names are converting into lower snake case; the 19 variables mentioned above are selected; and `entry` is converted into logical variable (`YES` to `TRUE`, `NO` to `FALSE`).
- Dimension: the dimension of resulting dataset is `r nrow(subway_df)` x `r ncol(subway_df)`.
